---
title: "Working with US Census Data"
author: "Bailey Morrison"
format: html
---

# Introduction 
## Load libraries
```{r}
#| warning: false
#| message: false

library(tidycensus)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(stringr)
```

## Getting Started
```{r}
#| warning: false
#| message: false
#| echo: false

# census_api_key("96620d628b0643d7d30d64598152c24aac3aa973", install = TRUE, overwrite = T)
```
# Quering data (focus on 2020 decennial data)
## Getting census data
```{r}
#| warning: false
#| message: false

pop_2020 <- get_decennial(
    geography = "state",
    variable = "P1_001N",
    year = 2020)

pop_2010 <- get_decennial(
  geography = "state", 
  variables = "P001001",
  year = 2010)
```

## Specific table
```{r}
#| warning: false
#| message: false

table_p2_2020 <- get_decennial(
    geography = "state",
    table = "P2",
    year = 2020)

unique(table_p2_2020$variable)
```

## Variable codes and definitions for  PL 94-171 
```{r}
#| warning: false
#| message: false

## for 2020
vars_pl_2020 <-  load_variables(2020, "pl")

## for 2010
vars_pl_2010 <-  load_variables(2010, "pl")
```


##  Demographic and Housing Characteristics File (DHC)
```{r}
#| warning: false
#| message: false

vars_dhc_2020 <-  load_variables(2020, "dhc")

## Note you have to specify the file with sumfile =
household_2020 <- get_decennial(
    geography = "state",
    variable = "H10_001N",
    year = 2020,
    sumfile = "dhc")
```

## Hispanic population the 6 counties around the Delta
```{r}
#| warning: false
#| message: false

delta_hispanic <- get_decennial(
  geography = "county",
  state = "CA",
  county = c("Alameda", "Contra Costa", "Sacramento", "San Joaquin", "Solano", "Yolo"),
  variables = "P2_002N",
  year = 2020)
```

# Quering for multiple variabels
## Concatenate function 
```{r}
#| warning: false
#| message: false

## Vector with race variables codes
race_vars <- c(
  Hispanic = "P2_002N",
  White = "P2_005N",
  Black = "P2_006N",
  Native = "P2_007N",
  Asian = "P2_008N",
  HIPI = "P2_009N") ## Native Hawaiian and other Pacific Islander

delta_race <- get_decennial(
  geography = "county",
  state = "CA",
  county = c("Alameda", "Contra Costa", "Sacramento", "San Joaquin", "Solano", "Yolo"),
  variables = race_vars,
  summary_var = "P2_001N",
  year = 2020)
```

# Getting Census Data ready for analysis
Notes on get_acs()
```{r}
## 1-year survey
median_income_1yr <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "CA",
  year = 2021,
  survey = "acs1")

## 5-year survey. Defaults to the 2017-2021 5-year ACS
median_income_5yr <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "CA")
```
## Load variables for “acs5”, year 2021.
```{r}
#| warning: false
#| message: false
## variables for 5-year 2017-2021 ACS
vars <- load_variables(2021, "acs5")
```

## Find code for total median gross rent.
Get acs data for median gross rent by county in California
```{r}
ca_rent <- get_acs(
  geography = "county",
  variables = "B25031_001",
  state = "CA",
  year = 2021)
```

## Visualize estimate by county using geom_point. (Hint: use y = reorder(NAME, estimate) to order counties by estimates)
```{r}
#| warning: false
#| message: false
ggplot(ca_rent, aes(x = estimate, y = reorder(NAME, estimate)))+
  geom_point()+
  theme_classic()
```

## Add the following layer, one by one, to your plot. Discuss with your neighbor what each line of code is doing.
```{r}
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 12


ggplot(ca_rent, aes(x = estimate, y = reorder(NAME, estimate)))+
  geom_point()+
  theme_classic()+
  labs(x = "Estimate ($)", y = "CA County")+
  geom_errorbar(aes(xmin = estimate - moe, xmax = estimate + moe),
                width = 0.5, linewidth = 0.5)+
  scale_x_continuous(labels = label_dollar())+
  scale_y_discrete(labels = function(x) stringr::str_remove(x, " County, California|, California"))

```

