---
title: "data-visualization"
format: html
---

# Setup

## Load Packages
```{r}
#| warning: false
#| message: false

library(readr)
library(dplyr)
library(tidyr)
library(forcats) # makes working with factors easier
library(ggplot2)
library(leaflet) # interactive maps
library(DT) # interactive tables
library(scales) # scale functions for visualization
library(janitor) # expedite cleaning and exploring data
library(viridis) # colorblind friendly color pallet
```

## Load Data
```{r}
#| warning: false
#| message: false
delta_visits <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.587.1&entityid=cda8c1384af0089b506d51ad8507641f") %>% 
    janitor::clean_names() ## Introducing this new package!
```

## Explore Data
```{r}
#| warning: false
#| message: false

## Check out column names
colnames(delta_visits)

## Peak at each column and class
glimpse(delta_visits)

## From when to when
range(delta_visits$date)

## First and last rows
head(delta_visits)
tail(delta_visits)

## Which time of day?
unique(delta_visits$time_of_day)
```

# Get Data Ready

## Make Long
```{r}
#| warning: false
#| message: false

visits_long <- delta_visits %>% 
    pivot_longer(cols = c("sm_boat", "med_boat", "lrg_boat", "bank_angler", "scientist", "cars"),
                 names_to = "visitor_type",
                 values_to = "quantity") %>%
    rename(restore_loc = eco_restore_approximate_location) %>% 
    select(-notes)

## Checking the outcome
head(visits_long)
```

## Plotting with `ggplot2`

```{r}
#| warning: false
#| message: false

ggplot(visits_long, aes(x = restore_loc, y = quantity))+
  geom_col()
```

```{r}
#| warning: false
#| message: false

daily_visits = visits_long %>%
  group_by(restore_loc, date, visitor_type) %>%
  summarise(daily_visits = sum(quantity))

head(daily_visits)
```

## Daily Visit locations
```{r}
daily_visits_loc <- visits_long %>%
    group_by(restore_loc, date, visitor_type) %>% 
    summarise(daily_visits = sum(quantity))
    
head(daily_visits_loc)
```

## ggplots
```{r}
## Option 1 - data and mapping called in the ggplot() function
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits))+
    geom_col()


## Option 2 - data called in ggplot function; mapping called in geom
ggplot(data = daily_visits_loc) +
    geom_col(aes(x = restore_loc, y = daily_visits))


## Option 3 - data and mapping called in geom
ggplot() +
    geom_col(data = daily_visits_loc,
             aes(x = restore_loc, y = daily_visits))
```

## Boxplot
```{r}
#| warning: false
#| message: false

daily_visits_loc %>%
    separate(date, c("year", "month", "day"), sep = "-") %>%
    filter(daily_visits < 30,
           visitor_type %in% c("sm_boat", "med_boat", "lrg_boat")) %>%
    ggplot(aes(x = visitor_type, y = daily_visits)) +
    geom_boxplot()
```

## Violin Plots
```{r}
daily_visits_loc %>%
    separate(date, c("year", "month", "day"), sep = "-") %>%
    filter(daily_visits < 30,
           visitor_type %in% c("sm_boat", "med_boat", "lrg_boat")) %>%
    ggplot(aes(x = visitor_type, y = daily_visits)) +
    geom_violin()
```

## Line and Point
```{r}
daily_visits_loc %>%
    filter(restore_loc == "Decker Island",
           visitor_type == "med_boat") %>%
    ggplot(aes(x = date, y = daily_visits)) +
    geom_line() +
    geom_point()
```

## Customize plots
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = "blue"))+
    geom_col()
```

## Change color
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits))+
    geom_col(fill = "blue")
```

## Color by Visitor type
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()
```

33 Setting ggplot themes
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during July 2017 and March 2018")+
    coord_flip()+
    theme_bw()
```

## Theme function for ggplot
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    theme_bw()+
    theme(legend.position = "bottom",
          axis.ticks.y = element_blank()) ## note we mention y-axis here
```

## Save ggplot theme
```{r}
my_theme <- theme_bw(base_size = 16) +
    theme(legend.position = "bottom",
          axis.ticks.y = element_blank())
```


## ggplot + myTheme
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    my_theme
```

## Scale y-axis
```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    my_theme
```

# Reorder Classes
## Create new variable
```{r}
daily_visits_totals <- daily_visits_loc %>% 
    group_by(restore_loc) %>%
    mutate(n = sum(daily_visits)) %>% 
    ungroup()

head(daily_visits_totals)
```

## Plot reordered data
```{r}
ggplot(data = daily_visits_totals,
       aes(x = fct_reorder(restore_loc, n), y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    my_theme
```

## Reverse the order
```{r}
ggplot(data = daily_visits_totals,
       aes(x = fct_reorder(restore_loc, desc(n)), y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    theme_bw()+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    my_theme
```

## Color with Viridis
```{r}
ggplot(data = daily_visits_totals,
       aes(x = fct_reorder(restore_loc, desc(n)), y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    theme_bw()+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    scale_fill_viridis_d()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    my_theme
```

## Save ggplot
```{r}
ggsave("figures/visit_restore_site_delta.jpg", width = 12, height = 6, units = "in")
```

# Facet Plots
## Create multiple plots
```{r}
facet_plot <- ggplot(data = daily_visits_totals,
       aes(x = visitor_type, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    theme_bw()+
    facet_wrap(~restore_loc,
               scales = "free_y",
               ncol = 5,
               nrow = 2)+
    scale_fill_viridis_d()+
    labs(x = "Type of visitor",
         y = "Number of Visits",
         title = "Total Number of Visits to Delta Restoration Areas",
         subtitle = "Sum of all visits during study period")+
    theme_bw()+
    theme(legend.position = "bottom",
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank())

facet_plot
```

## Save facet plots
```{r}
ggsave("figures/visit_restore_site_facet.jpg", facet_plot, width = 12, height = 8, units = "in")
```


# Tables with `DT`
## Create dataframe
```{r}
locations <- visits_long %>%
    distinct(restore_loc, .keep_all = T) %>%
    select(restore_loc, latitude, longitude)

head(locations)
```

## Display interactive table
```{r}
datatable(locations)
```

# Maps with `leaflet`
## Spatial interactive map
```{r}
leaflet(locations) %>%
    addTiles() %>%
    addMarkers(
        lng = ~ longitude,
        lat = ~ latitude,
        popup = ~ restore_loc
    )
```

## Import Web Map Service (WMS)
```{r}
leaflet(locations) %>%
    addWMSTiles(
        "https://basemap.nationalmap.gov/arcgis/services/USGSTopo/MapServer/WmsServer",
        layers = "0",
        options = WMSTileOptions(format = "image/png", transparent = TRUE)) %>%
    addCircleMarkers(
        lng = ~ longitude,
        lat = ~ latitude,
        popup = ~ restore_loc,
        radius = 5,
        # set fill properties
        fillColor = "salmon",
        fillOpacity = 1,
        # set stroke properties
        stroke = T,
        weight = 0.5,
        color = "white",
        opacity = 1)
```

## Base Maps
```{r}
leaflet(locations) %>%
    addWMSTiles(
        "https://basemap.nationalmap.gov/arcgis/services/USGSImageryTopo/MapServer/WmsServer",
        layers = "0",
        options = WMSTileOptions(format = "image/png", transparent = TRUE)) %>%
    addWMSTiles(
        "https://basemap.nationalmap.gov/arcgis/services/USGSHydroCached/MapServer/WmsServer",
        layers = "0",
        options = WMSTileOptions(format = "image/png", transparent = TRUE)) %>%
    addCircleMarkers(
        lng = ~ longitude,
        lat = ~ latitude,
        popup = ~ restore_loc,
        radius = 5,
        # set fill properties
        fillColor = "salmon",
        fillOpacity = 1,
        # set stroke properties
        stroke = T,
        weight = 0.5,
        color = "white",
        opacity = 1)
```

# Publish the Data Visualization lesson to your webpage




