---
title: "Working with Text Data"
author: "Bailey Morrison"
format: html
---

```{r}
library(gutenbergr) # access public domain texts from Project Gutenberg
library(tidytext) # text mining using tidy tools
library(dplyr) # wrangle data
library(ggplot2) # plot data
```

```{r}
gutenberg_works(title == "Dracula") # dracula text
```

Get the id number from the gutenberg_works() function so that you can download the text as a corpus using the function gutenberg_download(). Save the corpus to an object called {book-title}_corp. View the object - is the data in a tidy format?
```{r}
dracula_corp = gutenberg_download(345)
```

Tokenize the corpus data using unnest_tokens(). Take a look at the data - do we need every single token for our analysis?
```{r}
tidy_dracula = dracula_corp %>%
  unnest_tokens(word, text)
```


Remove “stop words” or words that can be safely removed or ignored without sacrificing the meaning of the sentence (e.g. “to”, “in”, “and”) using anti_join().

Take a look at the data - are you satisfied with your data? We won’t conduct any additional cleaning steps here, but consider how you would further clean the data.
```{r}
tidy_dracula = tidy_dracula %>%
  dplyr::anti_join(stop_words, by = "word")
```

Calculate the top 10 most frequent words using the functions count() and slice_max().
```{r}
count_dracula <- tidy_dracula %>%
    count(word) %>% 
    slice_max(n = 10, order_by = n)
```

```{r}
# bar plot
ggplot(data = count_dracula, aes(n, reorder(word, n))) +
  geom_col() +
    labs(x = "Count",
         y = "Token")
```

```{r}
# bar plot
ggplot(data = count_dracula, aes(n, reorder(word, n))) +
  geom_col(fill = "chartreuse4") +
    labs(x = "Count",
         y = "Token")+
  ggtitle("Top 10 Dracula Words")+
  theme_classic()+
  theme(plot.title = element_text(size = 20),
        axis.text.y = element_text(size = 12))
```

```{r}
# initial lollipop plot
ggplot(data = count_dracula, aes(x=word, y=n)) +
    geom_point() +
    geom_segment(aes(x=word, xend=word, y=0, yend=n)) +
    coord_flip() +
    labs(x = "Token",
         y = "Count")

```

```{r}

# ascending order pretty lollipop plot
ggplot(data = count_dracula, aes(x=reorder(word, n), y=n)) +
    geom_point(color="cyan4") +
    geom_segment(aes(x=word, xend=word, y=0, yend=n), color="cyan4") +
    coord_flip() +
    labs(title = "Top Ten Words in The Phantom of the Opera",
         x = NULL,
         y = "Count") +
    theme_minimal() +
    theme(
        panel.grid.major.y = element_blank()
    )
```

