---
title: "Literate_analysis"
author: "Bailey D Morrison"
format: html
editor: visual
---

## Introduction

### About this data
This data set contains North Pole Environmental Observatory Bottle Chemistry data. This data set includes diverse measurements of seawater chemistry along with supporting conductivity, temperature, and depth (CTD) data from the Arctic Ocean near the North Pole. Measurements were taken from sea ice platforms each April or May from 2000-2014.
[https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

### Setup
```{r}
#| message: false
require(readr)
```

### Read in data
```{r}
#| warning: false
#| message: false
bg_chem = read_csv("data/BGchem2008data.csv")
```

Exploring data
```{r}
#| eval: false
#| echo: false
colnames(bg_chem)

str(bg_chem)

head(bg_chem)

summary(bg_chem)

unique(bg_chem$Date)
```


## Analysis

### Calcualte summary stats
```{r}
#| warning: false
#| message: false
nitrate = mean(bg_chem$NO3, na.rm =T)
nitrite = mean(bg_chem$NO2, na.rm = T)
amm = mean(bg_chem$NH4, na.rm = T)
phos = mean(bg_chem$P, na.rm = T)
```

### Calculate mean Redfield Ratio
```{r}
#| warning: false
#| message: false
ratio = (nitrate + nitrite + amm)/phos
print(ratio)
```

### Plot Redfield Ratio
```{r}
plot(bg_chem$P, (bg_chem$NH4 + bg_chem$NO3 + bg_chem$NO2), xlab = "Phosphorus", ylab = "Nitrogen")
```


## Conclusion
Redfield Ratio for this data is `r round(ratio, 0)`.

