---
title: "Spatial Analysis"
author: "Bailey Morrison"
format: html
---

# About the data
## Setup
```{r}
# knb_url <- 'https://dev.nceas.ucsb.edu/knb/d1/mn/v2/object/urn%3Auuid%3Aaceaecb2-1ce0-4d41-a839-d3607d32bb58'
# download.file(url = knb_url, destfile = 'demo_data.zip')
# unzip('demo_data.zip', exdir = 'data')
# file.remove('demo_data.zip')
```

## Load Libraries
```{r}
#| warning: false
#| message: false

library(readr)
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
library(dplyr)
```

## xploring the data using  `plot()` and `st_crs()`
```{r}
#| warning: false
#| message: false

# read in shapefile using read_sf()
ak_regions <- read_sf("data/ak_regions_simp.shp")

# quick plot
plot(ak_regions)
```

```{r}
#| warning: false
#| message: false

class(ak_regions)
```

```{r}
#| warning: false
#| message: false

head(ak_regions)
```

```{r}
#| warning: false
#| message: false

glimpse(ak_regions)
```

## Coordinate Reference System (CRS)
```{r}
#| warning: false
#| message: false

st_crs(ak_regions)
```

```{r}
#| warning: false
#| message: false

ak_regions_3338 <- ak_regions %>%
    st_transform(crs = 3338)

st_crs(ak_regions_3338)
```

```{r}
#| warning: false
#| message: false

plot(ak_regions_3338)
```

# `sf` & the Tidyverse
## `select()'
```{r}
#| warning: false
#| message: false

# returns the names of all the columns in dataset
colnames(ak_regions_3338)
```

```{r}
#| warning: false
#| message: false

ak_regions_3338 %>%
    select(region)
```

## `filter()`
```{r}
#| warning: false
#| message: false

unique(ak_regions_3338$region)
```

```{r}
#| warning: false
#| message: false

ak_regions_3338 %>%
    filter(region == "Southeast")
```

# Spatial Joins
##  Read in alaska_population.csv using `read.csv()`
```{r}
#| warning: false
#| message: false

# read in population data
pop <- read_csv("data/alaska_population.csv")
```

## Turn `pop` into a spatial object
```{r}
#| warning: false
#| message: false

pop_4326 <- st_as_sf(pop,
                     coords = c('lng', 'lat'),
                     crs = 4326,
                     remove = F)

head(pop_4326)
```
## Example: Makes an error
```{r}
# #| warning: false
# #| message: false
# 
# pop_joined <- st_join(pop_4326, ak_regions_3338, join = st_within)
```

```{r}
#| warning: false
#| message: false

pop_3338 <- st_transform(pop_4326, crs = 3338)
```

```{r}
#| warning: false
#| message: false

pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

head(pop_joined)
```

## Calculate the total population by region using `group_by()` and `summarize()`
```{r}
#| warning: false
#| message: false

pop_region <- pop_joined %>%
    as.data.frame() %>%
    group_by(region) %>%
    summarise(total_pop = sum(population))

head(pop_region)
```

```{r}
#| warning: false
#| message: false

pop_region_3338 <- left_join(ak_regions_3338, pop_region, by = "region")

# plot to check
plot(pop_region_3338["total_pop"])
```

## Calculate the population by Alaska management area, as opposed to region
```{r}
#| warning: false
#| message: false

pop_mgmt_338 <- pop_region_3338 %>%
    group_by(mgmt_area) %>%
    summarize(total_pop = sum(total_pop))

plot(pop_mgmt_338["total_pop"])
```

## If we don’t want to combine geometries, we can specify do_union = F as an argument
```{r}
#| warning: false
#| message: false

pop_mgmt_3338 <- pop_region_3338 %>%
    group_by(mgmt_area) %>%
    summarize(total_pop = sum(total_pop), do_union = F)

plot(pop_mgmt_3338["total_pop"])
```

## Save the spatial object to a new file using` `write_sf()`
```{r}
#| warning: false
#| message: false

write_sf(pop_region_3338, "data/ak_regions_population.shp")
```

