---
title: "survey-workflow"
format: html
---

## Setup
```{r}
#| warning: false
#| message: false
library(qualtRics)
library(tidyr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(dplyr)
```


```{r}
#| warning: false
#| message: false
# qualtrics_api_credentials(api_key = "Zzuvz4t36VkB5akSB8W3Dryg5YgZ9lowoc3JEfEN", base_url = "ucsb.co1.qualtrics.com", install = TRUE, overwrite = T)
```

## Get all surveys
To get a list of all the surveys in your Qualtrics instance, use the all_surveys function.
```{r}
#| warning: false
#| message: false
surveys <- all_surveys()
kable(surveys) %>%
    kable_styling()
```

## Save survey id
```{r}
#| warning: false
#| message: false
i <- which(surveys$name == "Survey for Data Science Training")
id <- surveys$id[i]
```

## Get questions
You can retrieve a list of the questions the survey asked using the survey_questions function and the survey id.
```{r}
#| warning: false
#| message: false
questions <- survey_questions(id)
kable(questions) %>%
    kable_styling()
```

## Get survey results
To get the full survey results, run fetch_survey with the survey id.
```{r}
#| warning: false
#| message: false
survey_results <- fetch_survey(id)
```


## Question 2
```{r}
#| warning: false
#| message: false
q2 <- survey_results %>% 
    select(Q2) %>% 
    group_by(Q2) %>% 
    summarise(n = n())
```


```{r}
#| warning: false
#| message: false
kable(q2, col.names = c("How long have you been programming?",
                        "Number of responses")) %>%
    kable_styling()
```

## Question 3
```{r}
#| warning: false
#| message: false
q3 <- survey_results %>% 
    select(Q3) %>% 
    group_by(Q3) %>% 
    summarise(n = n())
```

```{r}
#| warning: false
#| message: false
ggplot(data = q3, 
       mapping = aes(x = Q3, y = n)) +
    geom_col() +
    labs(x = "What language do you currently use most frequently?", y = "Number of reponses") +
    theme_minimal()
```

```{r}
#| warning: false
#| message: false
q3_text <- survey_results %>% 
    select(Q3_7_TEXT) %>% 
    drop_na()

kable(q3_text, col.names = c("Other responses to 'What language do you currently use mose frequently?'")) %>% 
    kable_styling()
```

## Question 4
```{r}
#| warning: false
#| message: false
q4 <- survey_results %>% 
    select(Q4) %>% 
    rename(`What data science tool or language are you most excited to learn next?` = Q4) %>% 
    drop_na()

kable(q4, col.names = "What data science tool or language are you most excited to learn next?") %>% 
    kable_styling()
```


